% Pilot Resources
% Description
Pilot Resources identifies the type of resources that the \pilot system exposes,
e.g. compute, data, or networking
% Example

% Pilot Deployment
% Description
Pilot Deployment describes the modalities of scheduling, utilization, and
aggregation of pilots.
% Example
For example, to schedule a pilot onto a specific DCR, \pilot systems need to
acquire and process information about what type of container to use (e.g. job,
virtual machine), what type of scheduler the resource exposes, the amount of
cores and duration they can be requisitioned for, what filesystems can be
accessed, or how to communicate with the DCR.

% Workload Semantics
% Description
Semantically, a workload description contains all the information necessary for
it to be dispatched to an appropriate DCR and to be executed.
% Example
For example, type,
number, size, and duration of tasks alongside their grouping into stages or
their data dependences need to be known when deciding how many resources of a
specific type should be used to execute the given workload but also for how long
such resources should be available.

% Workload Binding
% Description
Executing a workload requires for its tasks to be bound to the resources. Both
the temporal and spatial dimensions of the binding operations are relevant for
the implementation of Task Dispatching.
% Example
Depending on the concurrency of a given workload, tasks could be dispatched to
one or more pilots for an efficient execution. Furthermore, tasks could be bound
to pilots before or after its instantiation, depending on resource availability
and scheduling decisions.

% Workload Execution
% Description
Workload Execution identifies the process of task dispatching and the modalities
with which the execution environment is set up for each task. The dispatching
process may also depend on the capabilities exposed by the target DCR and its
policies. As such, the execution of the workload may depend on how the \pilot
systems implement their interactions with the remote DCR.
% Example

